---
title: 'Assigment: Plot deconstruction'
author: "Fernando Cisneros & Maria Paula Muñoz"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## R Markdown

## Plot deconstruction: Plastic Ingestion in Post-hatchling Sea Turtles: Assessing a Major Threat in Florida Near Shore Waters

```{r}
library(readxl)
st <- read_excel("st.xlsx")
View(st)
```

# Transform to log 

```{r}
st$log_total <- log(st$total + 1)  # Adding 1 to avoid log(0)

```

```{r}
# Reorder the factor levels of 'type' based on 'total' values in descending order
st$type <- factor(st$type, levels = st$type[order(st$total, decreasing = TRUE)])
```




# Making the new graph

```{r}
library(tidyverse)
library(ggplot2)
ggplot(st, aes(x = type, y = log_total)) + 
  geom_bar(stat = "identity", fill = "black") +  xlab("Type of Plastic Ingested") +
  ylab("Logarithmic Total") + theme_minimal() 
```
```{r}
ggplot(st, aes(x = type, y = log_total)) + 
  geom_bar(stat = "identity", fill = "grey30", color = "black") +  # Set bars to grey with black borders
  xlab("Type of Plastic Ingested") +  # Adding x-axis label
  ylab("Logarithmic Total") +  # Adding y-axis label
  theme_classic() +  # Use a classic theme for a clean look
  theme(
    panel.grid.major = element_line(color = "grey80"),  # Major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.text = element_text(size = 12),  # Adjust axis text size
    axis.title = element_text(size = 14, face = "bold"),  # Adjust axis title size and style
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to the top of x-axis title
    axis.title.y = element_text(margin = margin(r = 10))   # Add margin to the right of y-axis title
  )
```




Comments para la grafica, el pie chart no da uncertanty o variabilidad. No nos deja ver SE, media y error bars. Explicar en la nueva data que seria bueno agregarlo en el bar chart. 
log-scale the Y axis. 



