---
title: "Assigment: End-to-end data analysis in R"
author: "Fernando Cisneros"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## End to end data analysis in R
## In this assignment, we will use real-world marine data collected as part of Queensland fisheries QFISH database to develop a report (either written on your website or via an R Markdown report embedded in your website) suitable for reporting to your manager, a community group or for your own research.

## Upload packages
```{r}
library(tidyverse)
```


## Shark Control Program
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
## the data was downloaded from the following link: https://qfish.fisheries.qld.gov.au/query/number-caught-by-area-calendar-year-and-species-group/table.
## the data selected was the shark control program by selecting the number caught by area, calendar year and species group offers a data table by species, year and region

## 
library(readxl)
sharks <- read_excel("~/Desktop/JCU/Data analysis/Mod 4 Personal/data/sharks.xlsx")
View(sharks)
```

## Reshaping the data
```{r message=TRUE, warning=TRUE, paged.print=TRUE}
sharks <- data_frame(
   Area = c("Bribie Island", "Bundaberg", "Cairns", "Capricorn Coast", "Gladstone", "Gold Coast", "Mackay", "Nth Strandbroke Island", "Rainbow beach", "Sunshine Coast North", "Sunshine Coast South", "Sunshine Coast & Bribie Island", "Townsville"), 
   `2001 Total` = c(0, 75, 100, 90, 32, 155, 172, 26, 44, 115, 0, 48, 108),
   `2002 Total` = c(0, 72, 70, 97, 20, 142, 112, 22, 47, 118, 0, 59, 68),
   `2003 Total` = c(0, 56, 87, 72, 34, 121, 98, 22, 29, 84, 0, 57, 72),
   `2004 Total` = c(0, 48, 84, 82, 59, 121, 135, 9, 54, 63, 0, 45, 84),
   `2005 Total` = c(0, 80, 88, 85, 46, 136, 83, 14, 109, 95, 0, 31, 101),
   `2006 Total` = c(0, 52, 56, 71, 56, 69, 77, 30, 112, 81, 0, 56, 111),
   `2007 Total` = c(0, 59, 97, 84, 49, 62, 75, 26, 89, 70, 0, 38, 79),
   `2008 Total` = c(0, 61, 74, 72, 52, 73, 90, 28, 80, 78, 0, 43, 111),
   `2009 Total` = c(0, 39, 75, 63, 63, 221, 69, 29, 89, 67, 0, 31, 135),
   `2010 Total` = c(0, 56, 95, 66, 57, 145, 78, 25, 79, 50, 0, 30, 149),
   `2011 Total` = c(0, 80, 113, 67, 76, 83, 69, 16, 78, 61, 0, 32, 185),
   `2012 Total` = c(0, 62, 114, 118, 67, 96, 107, 19, 72, 36, 0, 35, 183),
   `2013 Total` = c(0, 81, 83, 121, 60, 64, 123, 18, 92, 34, 0, 23, 114),
   `2014 Total` = c(0, 65, 71, 73, 50, 78, 121, 9, 50, 37, 0, 18, 120),
   `2015 Total` = c(0, 57, 116, 75, 46, 81, 132, 24, 67, 59, 0, 29, 156),
   `2016 Total` = c(0, 53, 53, 58, 52, 86, 68, 12, 55, 49, 0, 28, 117),
   `2017 Total` = c(1, 55, 47, 43, 45, 126, 58, 17, 66, 60, 11, 14, 112),
   `2018 Total` = c(2, 54, 43, 45, 36, 104, 41, 22, 59, 94, 24, 0, 120),
   `2019 Total` = c(1, 59, 43, 39, 43, 130, 45, 17, 71, 85, 18, 0, 91),
   `2020 Total` = c(1, 56, 63, 51, 30, 129, 73, 19, 75, 91, 15, 0, 143),
   `2021 Total` = c(2, 75, 59, 148, 37, 114, 95, 39, 96, 98, 36, 0, 160),
   `2022 Total` = c(4, 64, 62, 196, 49, 102, 62, 28, 81, 83, 37, 0, 133),
   `2023 Total` = c(4, 56, 66, 251, 93, 116, 76, 19, 67, 121, 40, 0, 187),
   `2024 Total` = c(0, 25, 66, 130, 60, 24, 19, 6, 26, 58, 14, 0, 67),
)


long_sharks <- sharks %>%
  pivot_longer(
    cols = Area,
    names_to = c("SpeciesGroup", "CalendarYear"),
    names_sep = "_",
    values_to = "Count"
  ) %>%
  mutate(
    CalendarYear = as.integer(CalendarYear)
  ) %>%
  filter(!is.na(Count))

```

##Inspect the data
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
head(long_sharks)
```

## PLotting the data
```{r}
ggplot(long_sharks, aes(x = CalendarYear, y = Count, color = "black")) +
  geom_line() +
  geom_point() +
  labs(
    title = "Total Counts Over Time by Area",
    x = "CalendarYear",
    y = "Total Count"
  ) +
  theme_minimal()
```

## Plotting a line graph
```{r}
# Reshaping data
long_sharks <- sharks %>%
  pivot_longer(
    cols = -Area,
    names_to = "Year",
    values_to = "Count"
  ) %>%
  mutate(Year = as.integer(str_remove(Year, " Total")))

# Plot
ggplot(long_sharks, aes(x = Year, y = Count, color = Area)) +
  geom_line() +
  labs(title = "Shark Counts by Area Over Years",
       x = "Year",
       y = "Shark Count",
       color = "Area") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1))
```

